---
title: "Transfers Research"
output:
  html_document:
    number_sections: no
    theme: paper
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'

---
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```{r setup, include=FALSE}

knitr::opts_chunk$set(comment=FALSE, message=FALSE, warning=FALSE, echo=FALSE
)

library(dplyr)
library(lubridate)
library(readxl)
library(tidyr)
library(ggplot2)
library(stringr)
library(htmltools)
library(tidyverse)
library(highcharter) 
#options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
library(ggalt)
library(cowplot)

```

```{r include=FALSE}

########################Cleaning the TRANSFER data###############################

library(writexl)

#import transfer excel that Shelby emailed
transfer <- read_excel("W:/Personal/Jen Z/UCLA Covid-19 Behind Bars/Data/Transfers Clean Data_from Shelby.xlsx",
                   sheet="Sheet3")


#############NOTE TO SELF: column for date is getting messed up when read into r
#cleaning data

transfer <- transfer[ -c(14:20) ] #keep columns of interest

#rename columns

names(transfer)[names(transfer) == "Notes...1"] <- "note1"
names(transfer)[names(transfer) == "Notes...12"] <- "note2"
names(transfer)[names(transfer) == "Date (Approx)"] <- "date_approx"
names(transfer)[names(transfer) == "From Facility"] <- "from_facility"
names(transfer)[names(transfer) == "From County"] <- "from_county"
names(transfer)[names(transfer) == "From State"] <- "from_state"
names(transfer)[names(transfer) == "To Facility"] <- "to_facility"
names(transfer)[names(transfer) == "#"] <- "people_count"
names(transfer)[names(transfer) == "To (State)"] <- "to_state"
names(transfer)[names(transfer) == "Reason"] <- "reason"
names(transfer)[names(transfer) == "Type"] <- "type"
names(transfer)[names(transfer) == "Facility/County"] <- "facility_county"

#now fix the date column. it is currently in the excel numeric format and needs to be in a date format we understand. 
transfer$date_approx<-format(as.Date(as.numeric(transfer$date_approx), origin = "1899-12-30"), "%m/%d/%y")

#now convert the column to date format

# transfer$date_approx<- ymd(transfer$date_approx)
# 
# #make the order of the date descending 
# transfer<-transfer[rev(order(transfer$date_approx)),]


```

```{r, include=FALSE}
################Lets clean up the data COVID CASE data###############################

library (readr)

#read in the data from github
urlfile="https://raw.githubusercontent.com/ivnagpal/ICE-COVID19/master/immdet_df.csv"

covid<-read_csv(url(urlfile))


#make character columns numeric -EXCEPT the first column with facility names which should be character

#covid<-covid %>% mutate_if(is.character,as.numeric)


covid[,2:684] <- sapply(covid[,2:684],as.numeric)


#replace the last '.' with a _ separater to separate the beginning DATE in the column name with the rest of the column name

substring(names(covid), 9, 9) <- "_"

#RENAME the 'Custody or Facility' column (first column) to something more normal

colnames(covid)[1] <- "Custody_or_Facility"


#lets convert the data from wide to long format
covid_long<-pivot_longer(covid, cols = -Custody_or_Facility, 
                         names_to = c("date","measure"), 
                         names_sep = "_", 
                         values_to = "value")


#now I want to add a new column that has the  7 day average across the metrics (covid cases, deaths, etc.)

covid_long<- covid_long%>%
                mutate(avg7day = zoo::rollmean(value, k = 7, fill = NA))


#now we will convert the date column from character to date for subsequent analyses
# covid_long<-(mutate(covid_long,date=ymd(date)))
# 
# covid_long$date <- ymd(covid_long$date)

#now I will write the data to an excel file so I can email it to Shelby
library("writexl")

# write_xlsx(covid_long, "W:/Personal/Jen Z/UCLA Covid-19 Behind Bars/Data/ICE_Covid19_downloaded 01-24-21.xlsx") #this is the file path for where I want the excel to go, anad the name I want to give the excel

```


```{r include=FALSE}

test<-transfer%>%
drop_na(date_approx)

#graph
hchart(transfer, "line",
         hcaes(x=date_approx, y=as.numeric(people_count), group=from_facility,
         tooltip = list(pointFormat = "Number of Transfers Made:
                        {point.people_count}<br>")))%>%

      hc_tooltip(crosshairs = TRUE)%>%
   hc_xAxis(title = list(text = "Approximate Date of Transfer"))%>%
  
  
  hc_yAxis(title = list(
    text= 'Number of Transfers')
    ,
    min=0,max=50
    )%>%
  
  hc_title(
    text = "Transfers Made",
    margin = 20,
    align = "left",
    style = list(useHTML = TRUE, fontWeight="bold")
    )






```


# Transfers Data 

## Transfers Made {.tabset .tabset-pills}

### Summary of Transfers Made by Facilities
```{r}
####NOTE NEED TO FIX and make it so 'NA' in the people_count column are treated as 0 and still aggregated


##########Summarize total SENT by facility############

sent<-transfer%>%
  group_by(from_facility)%>%
  summarize(total_sent=sum(as.numeric(people_count)))

#make the total  count column be in descending order
sent<-sent[order(sent$total_sent, decreasing = TRUE),]  


#make a prettier looking table

library(kableExtra)
kable(sent, col.names = c("Facility Name", "Total People Transferred Out"))%>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))%>%
   row_spec(1:5, bold = T, color = "black", background = "#E0F3F7")
  


```


### Facilities who make the most transfers (Top 5)
```{r}


top5 <- c("Buffalo Federal Detention Facility", "Irwin County Detention Center", "Pine Prairie ICE Processing Center", 
               "Stewart Detention Center", "South Louisiana Correctional Center")

sent_top5<-transfer%>%
  filter(from_facility %in% top5)%>%
  group_by(from_facility)%>%
  arrange(desc(people_count))



sent_top5 %>% 
  hchart('column', hcaes(x = from_facility, y = as.numeric(people_count), group=to_facility),
         tooltip = list(pointFormat = "Facility Transferred to: {point.to_facility}<br>
                        Total Number of People: {point.people_count}<br>
                        Date of Transfer: {point.date_approx}"))%>% 
    hc_tooltip(crosshairs = TRUE)%>%
  hc_xAxis(title = list(text = ""))%>%
    hc_yAxis(title = list(text = "Number of People Transferred"))%>%
  hc_title(
    text = "Facilities that Make the Most Transfers",
    margin = 20,
    align = "left",
    style = list(useHTML = TRUE, fontWeight="bold")
    )%>%
  hc_subtitle(text="Where do high-transfer facilities send inmates",
              align="left")




```

## Transfers Recieved {.tabset .tabset-pills}

### Summary of Transfers Recieved by Facilities

It is important to note that a majority of inmates who are being transferred are going to facilities we do not have information on (see rows in red)
```{r}

###############Summarize facilities who RECIEVED transfers#########################

####NOTE NEED TO FIX and make it so 'NA' in the people_count column are treated as 0 and still aggregated

recieved<-transfer%>%
  group_by(to_facility)%>%
  summarize(total_recieved=sum(as.numeric(people_count)))

#make the total  count column be in descending order
recieved<-recieved[order(recieved$total_recieved, decreasing = TRUE),]  


#make a prettier looking table

library(kableExtra)

kable(recieved, col.names = c("Facility Name", "Total Transfers Recieved"))%>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))%>%
   row_spec(2:3, bold = T, color = "black", background = "#E0F3F7")%>%
     row_spec(5, bold = T, color = "black", background = "#E0F3F7")%>%
       row_spec(1, bold = T, color = "black", background = "#F7E4E0")%>%
       row_spec(4, bold = T, color = "black", background = "#F7E4E0")

```

### Facilities who recieved the most transfers (Top 5)
```{r}
top3 <- c("Prairieland Detention Facility",	
"Stewart Detention Center",	
"Broward Transitional Center")


recieve_top3<-transfer%>%
  filter(to_facility %in% top3)%>%
  group_by(to_facility)%>%
  arrange(desc(people_count))

#graph

recieve_top3 %>% 
  hchart('column', hcaes(x = to_facility, y = as.numeric(people_count), group=from_facility),
         tooltip = list(pointFormat = "Facility Transferred to: {point.to_facility}<br>
                        Total Number of People: {point.people_count}<br>
                        Date Transfer Recieved: {point.date_approx:%H:%M}"))%>%
  hc_tooltip(crosshairs = TRUE)%>%
  hc_xAxis(title = list(text = ""))%>%
    hc_yAxis(title = list(text = "Number of People Transferred"))%>%
  hc_title(
    text = "Facilities that Recieve Most Transfers",
    margin = 20,
    align = "left",
    style = list(useHTML = TRUE, fontWeight="bold")
    )%>%
  hc_subtitle(text="Where to facilities recieve transfers from",
              align="left")


```


# Active Covid Cases Data from Top Facilities {.tabset .tabset-pills}

## Top 5 Facilities Covid Cases

```{r}
library(htmltools)
library(htmlwidgets)
library(highcharter)

top5<-covid_long%>%
  filter(grepl('Total confirmed COVID-19 cases', measure))%>%
  filter(grepl("Broward|Glades|Krome|Stewart|Wakulla", Custody_or_Facility))

top5$date<-mdy(top5$date)


#graph --NOTE: do we want to plot the '7 day average' or the actual cases' ?

  highcharter::hchart(top5, hcaes(x='date', y='value', group='Custody_or_Facility'), type="line")%>%
    hc_yAxis(title=list(text="Number of confirmed cases"))%>%
    hc_xAxis(type = 'datetime', title=list(text=""))
  
  # highcharter::hchart(state_race_subset, hcaes(x = 'date', y = 'avg_7day_cases_rate_ratio', group = 'race_ethn'), type = "line",
  #               tooltip = list(pointFormat = "{point.race_ethn} <br> The 7-day average case rate among {point.race_ethn}s was <br> {point.avg_7day_cases_rate_ratio} times the case rate of the lowest group on this day <br> {point.avg_new_7day_cases_rate} new cases per 100,000 in the past 7 days")) %>%
  # hc_yAxis(title = list(text = "X times the 7-day average of the lowest group")) %>%
  # # hc_xAxis(title = list(text = "Race/Ethnicity")) %>%
  #   hc_tooltip(crosshairs = TRUE) %>%
  # hc_xAxis(type = 'datetime',plotLines=plotLines)

  
             
```

## Prairieland Detention Facility 

```{r}


prarieland<-covid_long%>%
  #filter(grepl('Total confirmed COVID-19 cases', measure))%>% ###NOTE we can choose if we only want to look at cases or other metrics like death 
  filter(grepl("Prairieland", Custody_or_Facility))

#convert data column

prarieland$date<-mdy(prarieland$date)


#grab transfer-recieved dates and make into a data frame

datevector<-c("10/09/20", "10/08/20", "10/07/20", "10/07/20", "10/05/20", "8/26/20", "8/25/20", "4/11/20")

datevector<-data.frame(datevector)

datevector$datevector <- mdy(datevector$datevector)


#create a plotline function for all the transfer dates

plotLines <- list();
for (i in 1:nrow(datevector)){
  plotLines[[i]] <- list(
    color = "#FF0000",
    width = 0.5,
    label=list(text='Transfer Recieved'),
    value = datetime_to_timestamp(as.Date(datevector[i,], tz = 'UTC')))
}



# graph

  highcharter::hchart(prarieland, "line", hcaes(x=date, y=value, group=measure))%>%
    hc_yAxis(title=list(text="Number of confirmed cases"))%>%
  hc_xAxis(type = 'datetime',plotLines=plotLines)

```

## Stewart Detention Center

```{r}


stewart<-covid_long%>%
  #filter(grepl('Total confirmed COVID-19 cases', measure))%>%
  filter(grepl("Stewart", Custody_or_Facility))

#convert data column
stewart$date<-mdy(stewart$date)


#grab transfer-recieved dates and make into a data frame

datevector<-c("10/08/20", "09/08/20", "08/31/20", "08/04/20", "06/15/20", "06/11/20", "05/26/20", "05/15/20", "05/14/20")

datevector<-data.frame(datevector)

datevector$datevector <- mdy(datevector$datevector)


#create a plotline function for all the transfer dates

plotLines <- list();
for (i in 1:nrow(datevector)){
  plotLines[[i]] <- list(
    color = "#FF0000",
    width = 0.5,
    label=list(text='Transfer Recieved'),
    value = datetime_to_timestamp(as.Date(datevector[i,], tz = 'UTC')))
}



# graph

  highcharter::hchart(stewart, "line", hcaes(x=date, y=value, group=measure))%>%
    hc_yAxis(title=list(text="Number of confirmed cases"))%>%
  hc_xAxis(type = 'datetime',plotLines=plotLines)
  

```

## Broward Detention Center

```{r}


broward<-covid_long%>%
  #filter(grepl('Total confirmed COVID-19 cases', measure))%>%
  filter(grepl("Broward", Custody_or_Facility))

#convert data column
broward$date<-mdy(broward$date)


#grab transfer-recieved dates and make into a data frame

datevector<-c("02/20/10", "12/20/06")

datevector<-data.frame(datevector)

datevector$datevector <- mdy(datevector$datevector)


#create a plotline function for all the transfer dates

plotLines <- list();
for (i in 1:nrow(datevector)){
  plotLines[[i]] <- list(
    color = "#FF0000",
    width = 0.5,
    label=list(text='Transfer Recieved'),
    value = datetime_to_timestamp(as.Date(datevector[i,], tz = 'UTC')))
}



# graph

  highcharter::hchart(broward, "line", hcaes(x=date, y=value, group=measure))%>%
    hc_yAxis(title=list(text="Number of confirmed cases"))%>%
  hc_xAxis(type = 'datetime',plotLines=plotLines)
  

```